---
title: "Mathematical Optimization for Decision Making"
author: "Yuanzhe(Roger) Li"
date: "12/03/2019"
output:
  beamer_presentation: 
    colortheme: seagull
    fonttheme: professionalfonts
    theme: Rochester
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
options(tinytex.verbose = TRUE)
library(dplyr)
library(ggplot2)
library(ompr)
library(ompr.roi)
library(ROI.plugin.glpk)
library(knitr)
library(tidyr)
library(ggrepel)
library(magrittr)
```

## Overview

- Basics of Linear Optimization
- Mixed-integer programming (MIP)
- MIP examples
  - Traveling salesman problem
  - Job scheduling
  - Resource allocation

## Basics of Linear Programming (LP)


## Mixed-Integer Programming (MIP)

$$\begin{aligned}
\max \quad & f(x)\\
s.t. \quad & g(x) \leq 0 \\
& x \in \mathbb{R} \times \mathbb{Z}
\end{aligned}$$

## Decisions

## Example 1 - traveling salesman problem (TSP)

- Problem definition (from [wikipedia](https://en.wikipedia.org/wiki/Travelling_salesman_problem))


*The travelling salesman problem (TSP) asks the following question: Given a list of cities and the distances between each pair of cities, what is the shortest possible route that visits each city exactly once and returns to the origin city?*


## Example of a random path
```{r tsp_setup, fig.width=3.8, fig.height=2.2, fig.align='center'}
# number of cities
n <- 10
# boundary of euclidean space
max_x <- 500
max_y <- 250
set.seed(9102)
cities <- data.frame(
  id = 1:n
  ,x = runif(n, max = max_x)
  ,y = runif(n, max = max_y)
)
# Add distances and dist_fun
distance <- as.matrix(
  stats::dist(select(cities, x, y)
  ,diag = TRUE
  ,upper = TRUE)
  )
dist_fun <- function(i, j) {
  vapply(seq_along(i), function(k) distance[i[k], j[k]], numeric(1L))
}
# Draw an example route
route <- c(5,1,9,3,10,4,6,2,8,7,5)
random_path <- data.frame(cid = route) %>%
  dplyr::left_join(cities, by = c("cid"="id")) %>%
  dplyr::mutate(
    origin=cid
    ,destination=lead(origin)
    ,Distance = dist_fun(origin, destination)
  )
total_dist <- sum(random_path$Distance, na.rm=TRUE)
# Plot the example cities and route
ggplot(cities, aes(x, y),label = id) + 
  geom_point() + 
  geom_text(aes(label=id),color = "blue",hjust=1.5, vjust=0.5, size =3) + 
  annotate(
    "text"
    ,label=paste0("Random path with total distance = ", round(total_dist,2))
    ,x= 250, y=225, color = "blue", size=2.5
    ) +
  geom_path(data = random_path, aes(color = Distance))+ 
  xlab("x coordinate") + 
  ylab("y coordinate") + 
  xlim(0,500) + ylim(0,250) + 
  ggtitle("TSP example with 10 cities") +
  theme_light() +
  theme( plot.title = element_text(size = 10, face = "bold"))
```
- Finding the optimal path by brute force takes $O(n!)$

## Formulate TSP as a MIP

\begin{block}{Miller-Tucker-Zemlin formulation}
  \begin{itemize}
    \item Label the cities with $1,...,n$ and define:
      $$
        x_{ij}= 
        \begin{cases}
            1& \text{if a path goes from city $i$ to city $j$}\\
            0              & \text{otherwise}
        \end{cases}
      $$
    \item Use another variable $u_i$ to denote the ordering of the cities.
  \end{itemize}
\end{block}


## Formulate TSP as a MIP (continued)
  $$\begin{aligned}
  \max \quad & \sum_{i=1}^{n} \sum_{j\neq i, j=1}^{n}d_{ij}x_{ij}\\
  s.t. \quad & x_{ij} \in \{0, 1\}               & i,j = 1,..., n\\
             & u_i \in \mathbf{Z}                & i = 2,..., n \\
             & \sum_{i=1, i\neq j}^{n}x_{ij} = 1 & j = 1,...,n \\
             & \sum_{j=1, j\neq i}^{n}x_{ij} = 1 & i = 1,...,n \\
             & u_i-u_j + nx_{ij} \leq n-1        & 2 \leq i\neq j \leq n\\ 
             & 0 \leq u_i \leq n-1               & 2 \leq i \leq n
  \end{aligned}$$

## MIP solution to TSP
```{r tsp_mip, include=FALSE}
tsp_model <- ompr::MIPModel() %>%
  # we create a variable that is 1 iff we travel from city i to j
  add_variable(x[i, j], i = 1:n, j = 1:n, 
               type = "integer", lb = 0, ub = 1) %>%
  
  # a helper variable for the MTZ formulation of the tsp
  add_variable(u[i], i = 1:n, lb = 1, ub = n) %>% 
  
  # minimize travel distance
  set_objective(sum_expr(dist_fun(i, j) * x[i, j], i = 1:n, j = 1:n), "min") %>%
  
  # you cannot go to the same city
  set_bounds(x[i, i], ub = 0, i = 1:n) %>%
  
  # leave each city
  add_constraint(sum_expr(x[i, j], j = 1:n) == 1, i = 1:n) %>%
  #
  # visit each city
  add_constraint(sum_expr(x[i, j], i = 1:n) == 1, j = 1:n) %>%
  
  # ensure no subtours (arc constraints)
  add_constraint(u[i] >= 2, i = 2:n) %>% 
  add_constraint(u[i] - u[j] + 1 <= (n - 1) * (1 - x[i, j]), i = 2:n, j = 2:n)
tsp_result <- solve_model(tsp_model, with_ROI(solver = "glpk", verbose = TRUE))
```
- We use the [GLPK](https://www.gnu.org/software/glpk/) solver to optimize the MIP formulated above.

```{r tsp_result}
tsp_sol <- ompr::get_solution(tsp_result, x[i, j]) %>% 
  dplyr::filter(value > 0)
opt_route <- c(1, 5, 9, 10, 3, 4 ,2, 8, 6, 7, 1)
opt_path <- data.frame(cid = opt_route) %>%
  dplyr::left_join(cities, by = c("cid"="id")) %>%
  dplyr::mutate(
    origin=cid
    ,destination=lead(origin)
    ,Distance = dist_fun(origin, destination)
  )
opt_dist <- sum(opt_path$Distance, na.rm=TRUE)
opt_path %>% 
  mutate(trip_id = 1:nrow(.)) %>%
  select(trip_id, origin, destination, Distance) %>%
  head(10) %>% 
  kable(digits=3, col.names = c("trip_id", "from ($i$)", "to ($j$)", "distance"))
```

## Visualize the optimal path to the TSP
```{r tsp_visual, fig.width=3.8, fig.height=2.8, fig.align='center'}
ggplot(cities, aes(x, y),label = id) + 
  geom_point() + 
  geom_text(aes(label=id),color = "blue",hjust=1.5, vjust=0.5, size =3) + 
  annotate(
    "text"
    ,label=paste0("Optimal path with total distance = ", round(opt_dist,2))
    ,x= 250, y=225, color = "blue", size=2.5
    ) +
  geom_path(data = opt_path)+ 
  xlab("x coordinate") + 
  ylab("y coordinate") + 
  xlim(0,500) + ylim(0,250) + 
  ggtitle("Optimal path to the TSP") +
  theme_light() +
  theme( plot.title = element_text(size = 10, face = "bold"))
```



## Example 2 - facility location problem (FLP)
- [Wikipedia definition](https://en.wikipedia.org/wiki/Facility_location_problem): *A branch of operations research concerned with the optimal placement of facilities to minimize transportation cost while considering other factors.*

- We consider a simple warehouse allocation problem, where facilities can be built in $m$ potential locations to serve $n$ customers. The goal is to minimize the total transportation cost and fixed facility cost.


## Visualization of a FLP
```{r flp_setup, fig.width=4, fig.height=2.2, fig.align='center'}
set.seed(1234)
# customers
grid_size <- 1000
n <- 100
customer_locations <- data.frame(
  id = 1:n,
  x = round(runif(n) * grid_size),
  y = round(runif(n) * grid_size)
)

# potential locations and fixed cost
m <- 20
warehouse_locations <- data.frame(
  id = 1:m,
  x = round(runif(m) * grid_size),
  y = round(runif(m) * grid_size)
)
fixedcost <- round(rnorm(m, mean = grid_size * 10, sd = grid_size * 5))


# transportation cost
transportcost <- function(i, j) {
  customer <- customer_locations[i, ]
  warehouse <- warehouse_locations[j, ]
  round(sqrt((customer$x - warehouse$x)^2 + (customer$y - warehouse$y)^2))
}

p <- ggplot(customer_locations, aes(x, y)) + 
  geom_point(alpha = 0.7, size=1) + 
  geom_point(data = warehouse_locations, color = "red", alpha = 0.5, shape = 15) +
  scale_x_continuous(limits = c(0, grid_size)) +
  scale_y_continuous(limits = c(0, grid_size)) +
  ggtitle("Warehouse location problem (n=100, m=20)") + 
  theme_light() +
  theme(axis.title = element_blank(), 
        axis.ticks = element_blank(), 
        axis.text = element_blank(), panel.grid = element_blank(),
        plot.title = element_text(size = 10, face = "bold"))
plot(p)
```
Black dots represent customers. Red triangles show potential locations to build warehouses.

## Formulate FLP as MIP
\begin{block}{Warehouse allocation problem}
  \begin{itemize}
    \item Let $I=\{1,...,n\}$ be the set of customers, $J=\{1,...,m\}$ the set of potential locations.
    \item Define two set of binary variables:
       $$ y_{j}= 
        \begin{cases}
            1& \text{a warehouse is built at location $j$}\\
            0              & \text{otherwise}
        \end{cases}$$
        and 
      $$ x_{ij}= 
        \begin{cases}
            1 & \text{if customer $i$ will be served by warehouse at $j$}\\
            0              & \text{otherwise}
        \end{cases}$$
  \end{itemize}
\end{block}


## Formulate FLP as a MIP


$$\begin{aligned}
  \max \quad & \sum_{i=1}^{n} \sum_{j}^{m}\text{transportation\_cost}_{ij}\cdot x_{ij}+ \sum_{j=1}^{m}\text{fixed\_cost}_{j}\cdot y_{j} \\
  s.t. \quad & \sum_{j}^{m} x_{ij} = 1,  \quad\forall i = 1,..., n \\
             & x_{ij} \leq y_j,  \qquad\forall          i = 1,...,n \quad j = 1,..., m \\
             & x_{ij} \in \{0, 1\},\quad\forall      i = 1,...,n \quad j = 1,..., m \\
             & y_j \in \{0, 1\},  \quad\forall         j = 1,..., m 
  \end{aligned}$$



## MIP solution to the FLP
```{r flp_model, include=FALSE}
# Construct model
model <- MIPModel() %>%
  # 1 iff i gets assigned to warehouse j
  add_variable(x[i, j], i = 1:n, j = 1:m, type = "binary") %>%
  
  # 1 iff warehouse j is built
  add_variable(y[j], j = 1:m, type = "binary") %>%
  
  # maximize the preferences
  set_objective(sum_expr(transportcost(i, j) * x[i, j], i = 1:n, j = 1:m) + 
                  sum_expr(fixedcost[j] * y[j], j = 1:m), "min") %>%
  
  # every customer needs to be assigned to a warehouse
  add_constraint(sum_expr(x[i, j], j = 1:m) == 1, i = 1:n) %>% 
  
  # if a customer is assigned to a warehouse, then this warehouse must be built
  add_constraint(x[i,j] <= y[j], i = 1:n, j = 1:m)

# Solve the model
result <- solve_model(model, with_ROI(solver = "glpk", verbose = TRUE))
```


```{r flp_solution, fig.width=4, fig.height=2.8, fig.align='center'}
suppressPackageStartupMessages(library(dplyr))
matching <- result %>% 
  get_solution(x[i,j]) %>%
  filter(value > .9) %>%  
  select(i, j)

plot_assignment <- matching %>% 
  inner_join(customer_locations, by = c("i" = "id")) %>% 
  inner_join(warehouse_locations, by = c("j" = "id"))
customer_count <- matching %>% group_by(j) %>% summarise(n = n()) %>% rename(id = j)
plot_warehouses <- warehouse_locations %>% 
  mutate(costs = fixedcost) %>% 
  inner_join(customer_count, by = "id") %>% 
  filter(id %in% unique(matching$j))
p + 
  geom_segment(data = plot_assignment, aes(x = x.y, y = y.y, xend = x.x, yend = y.x), alpha = 0.8, color="blue") + 
  geom_point(data  = plot_warehouses, color = "red", size = 3, shape = 15) +
  ggrepel::geom_label_repel(data  = plot_warehouses, 
                            aes(label = paste0("fixed costs:", costs, " - # of customers: ", n)), 
                            size = 2, nudge_y = 20) + 
  ggtitle("Optimal warehouse locations and customer assignment")
```
