---
title: "The Shortest Primer on Mathematical Optimization"
author: "Yuanzhe(Roger) Li"
date: "10/29/2019"
output:
  beamer_presentation: 
    colortheme: seagull
    fonttheme: professionalfonts
    theme: Rochester
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
options(tinytex.verbose = TRUE)
library(dplyr)
library(ggplot2)
library(ompr)
library(ompr.roi)
library(ROI.plugin.glpk)
```

## R Markdown

This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

## Overview

- Mathematical Optimization
- Mixed-Integer Programs
  - Examples 

## Overview of Mathematical Optimization


## Mixed-Integer Programs

$$\begin{aligned}
\max \quad & f(x)\\
s.t. \quad & g(x) \leq 0 \\
& x \in \mathbb{R} \times \mathbb{Z}
\end{aligned}$$



## Portfolio Optimization


## Traveling Salesman Problem
```{r TSP}
```
